Temporality is Important for Networks of Hospital Patient Transfers

### 2025-04-07 reports/_model.qmd
restruct model part

---
title: "Section: Model of Stochastic Processes and Network Models"
filters:
  - tikz 
format: html
execute:
  echo: false
---

Notations

- $u,v$ for locations; discrete, usually subscript
- $s,t$ for times; continuous, usually function inputs/variables
- $T, \omega, \Omega$ for characteristic times
- $N$ as a size variable
- $x, y, z$ are state variables; often functions of location and time

@@

## General Sketch

@@

  \node[draw, squarx] (A) {$u$};
  \node (X) [right=of A] {};
  \node[draw, squarx] (B) [right=of X] {$v$};
  \node[draw, house] (H) [below=of X] {};
  \node[draw, circle] (D) [above=of X] {$\emptyset$};

@@

Sketch of processes modelled
:::

@@

Note that

$$ \lambda(U \to V) = \lambda $$

represents that in some time step $\Delta t$, the number of individuals that move from state $U$ to state $V$ is distributed as

$$ \mathrm{Poisson}(\lambda x_U \Delta t) $$.

@@

For both the naïve and improved static network models, we have that the observations of $w$ are taken over a time period of $T_\Sigma$.

In the naïve static model , we have that:

$$d_{uv}(s,t) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv} = 0$$
$$\rho_{uv}(s,t) = \rho_{uv} = 0$$

That is, all movements are instantaneous, and the rate of movement is the mean rate of movement over the entire observation period.

@@

For the improved static model, we choose some threshold value $\omega$ that delineates direct and indirect transfers.

We will have:

@@

For the snapshot model, we choose a snapshot duration $\omega$. This defines the threshold duration of an indirect transfer.

This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit to their next hospital.
This alters the indirect transfer processes to

@@

where $\delta(\cdot)$ is the Dirac delta function.

Alongside this, we also compute from the data, the number of patients "at home" at a given time, in order to inform the denominator for the process from $z'$ to $z$. We denote this quantity $H_{uv}(t)$

We will have that:

@@

For the temporal network, we make a choice of the time discretisation $\omega$. That is, events that occur within a time window $[\tau, \tau+\omega)$ are collapsed and aggregated to occur at time $\tau$. Over the observation (and simulation) period, we will have a series of time window boundaries $\{\tau_0, \tau_1, \dots\}$. We denote the smallest value in that set larger than a given time $t$ to be $\ceil{t}_\omega$

For the readmission process, we assume that individuals that would readmit within a given time window $[\tau, \tau+\omega)$ do so uniformly.

This yields:

@@

The readmission process thus has a uniform hazard.

>>>>

<!-- Notations

- $u,v$ for locations; discrete, usually subscript
- $s,t$ for times; continuous, usually function inputs/variables
- $T, \omega, \Omega$ for characteristic times
- $N$ as a size variable
- $x, y, z$ are state variables; often functions of location and time -->

@@

@@

  \node[draw, squarx] (A) {$u$};
  \node (X) [right=of A] {};
  \node[draw, squarx] (B) [right=of X] {$v$};
  \node[draw, house] (H) [below=of X] {$h_{uv}$};
  \node[draw, circle] (D) [above=of X] {$\emptyset$};

@@

Sketch of the generalised model with processes of 1. permanent discharge ($u \to \emptyset$), 2. direct transfer ($u \to v$), 3. return home ($u \to h_{uv}$, 4. readmission $h_{uv} \to v$).
:::

@@

We note that as a consequence of the model structure, the hazards can also be interpreted as modelling the hodling time an individual patient spends in each state.

@@

We can begin with a simple naive static model. This model will have the edge weights between healthcare facilities that represent the number of transfers from the source node to the target node over a certain period of observation time ($T_\Sigma$), regardless of how long the patient is absent from the healthcare facilities between initial discharge and subsequent readmission.

With reference to the general model above, the hazards of the  naïve static model are:
$$d_{uv}(s,t) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv} = 0$$
$$\rho_{uv}(s,t) = \rho_{uv} = 0$$

That is, all movements are instantaneous, and the rate of movement is the mean rate of movement over the entire observation period. This is an oversimplification of the work done in @donker, but represents a "worst" case model of temporal fidelity that can be benchmarked against.

@@

We can introduce the concept of "indirectness" by allowing individuals to return home, in an "improved" static model.
In this model, we choose some threshold value $\omega$ that delineates direct transfers that occur "instantaneously" and indirect transfers that require a patient to first move to an intermediary "home" state before readmitting.

In comparison to the naive model, we will have non-zero hazards $\eta$ and $\rho$:

@@

Of course, as the hospital system changes over time, the rate of patient transfers, and thus the hazards represented in our model should also change.
We see evidence of this in the overall rates of movement in the network over time

.... need a plot here of total direct/direct over time

We approach this using static network snapshots of the hospital system over time.

For the snapshot model, we choose a snapshot duration $\omega$. This defines the threshold duration of an indirect transfer.
This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit to their next hospital.
This alters the indirect transfer processes to

@@

where $\delta(\cdot)$ is the Dirac delta function.
Alongside this, we also compute from the data, the number of patients "at home" at a given time, in order to inform the denominator for the process from $z'$ to $z$. We denote this quantity $H_{uv}(t)$
We will have that:

@@

For the temporal network, we make a choice of the time discretisation $\omega$. That is, events that occur within a time window $[\tau, \tau+\omega)$ are collapsed and aggregated to occur at time $\tau$. Over the observation (and simulation) period, we will have a series of time window boundaries $\{\tau_0, \tau_1, \dots\}$. We denote the smallest value in that set larger than a given time $t$ to be $\ceil{t}_\omega$
For the readmission process, we assume that individuals that would readmit within a given time window $[\tau, \tau+\omega)$ do so uniformly. 
This yields:

@@

### 2025-04-07 reports/report.qmd
restruct model part

@@

## Movement Simulation Studies

>>>>

## Network Models for Temporal Fidelity

Static networks provide a very interpretable model of influence of nodes in the system: network measures allow for some semblance of understanding of aggregate aspects of the network and its components. However, dynamic models derived from those static networks miss important dynamics and temporal qualities of the system they are modelling.
It is well-known that temporal networks exhibit different behaviour to analogous static networks: they have lower connectivity ....

In the context of the spread of AMR between healthcare facilities, these differences can lead to misleading results about the dynamics of the [disease], and related control mechanisms and policies.
Using the detailed linkage data from the VAED, we can constrct a relatively detailed temporal network, which can then be projected in various network representations with different temporal fidelity.
We roughly consider three aspects of the dynamics:

1. the representation of the time patients spend outside the healthcare system (i.e. at home)
1. the time-varying number of transfers (direct and indirect) between facililties
1. the distribution of the inter-admission times for patients that are discharged and subsequently readmit to the system

{{< include _model.qmd >}}

@@

## Movement Simulation Studies
We simulate the movement of individuals without infection dynamics (for example, in the case of a noncommunicable disease) on each of the different types of network.
$X_0 = 5$ individuals are seeded in a single healthcare facility at the start of each simulation, and move through the network for $T = 8$years.
